---
title: "Anemia and Hb in Shuar"
author: "Alicia DeLouize"
date: "November 16, 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Revisions for first round of Revise and Resubmit to American Journal of Human Biology marked with (RandR2)

# Data
```{r}
set.seed(2323)

# Load packages 
library(readxl)
library(psych)
library(lme4)
library(tidyverse)
library(jtools)
library(scales)
library(reshape2)
library(ggpubr)
library(performance)
library(moments)
library(mgcv)
library(voxel)
library(lsr)
library(arsenal)

# attach data 
load("Hb.RData")
```

# Data Manipulation - Create Anemia Groups
```{r}
Hb$AnemiaGroups <- Hb$AgeGroups
Hb$AnemiaGroups[which(Hb$AnemiaGroups == 3 & Hb$Sex == 0)] <- 5
Hb$AnemiaGroups[which(Hb$AnemiaGroups == 3 & Hb$Sex == 1 & Hb$Pregnant == 0)] <- 3
Hb$AnemiaGroups[which(Hb$AnemiaGroups == 3 & Hb$Sex == 1 & Hb$Pregnant == 1)] <- 4
describe(Hb$AnemiaGroups)
describe(Hb$AgeGroups)
```

# Data Manipulation - Create Anemia 
```{r}
Hb$Anemia <- 0
Hb$Anemia <- ifelse(Hb$AnemiaGroups == 0 & Hb$Hb_AltAdj < 11.0, 1, 
ifelse(Hb$AnemiaGroups == 1 & Hb$Hb_AltAdj < 11.5, 1, 
ifelse(Hb$AnemiaGroups == 2 & Hb$Hb_AltAdj < 12.0, 1, 
ifelse(Hb$AnemiaGroups == 3 & Hb$Hb_AltAdj < 12.0, 1, 
ifelse(Hb$AnemiaGroups == 4 & Hb$Hb_AltAdj < 11.0, 1, 
ifelse(Hb$AnemiaGroups == 5 & Hb$Hb_AltAdj < 13.0, 1, 0))))))

```

# Data Management - Anemia Mild Moderate Severe  
```{r}
# Creating Mild Category
Hb$AnemiaMild <- 0
Hb$AnemiaMild <- ifelse(Hb$AnemiaGroups == 0 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 10.00, 1, 
ifelse(Hb$AnemiaGroups == 1 & Hb$Hb_AltAdj < 11.5 & Hb$Hb_AltAdj >= 11.00, 1, 
ifelse(Hb$AnemiaGroups == 2 & Hb$Hb_AltAdj < 12.0 & Hb$Hb_AltAdj >= 11.00, 1, 
ifelse(Hb$AnemiaGroups == 3 & Hb$Hb_AltAdj < 12.0 & Hb$Hb_AltAdj >= 11.00, 1, 
ifelse(Hb$AnemiaGroups == 4 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 10.00, 1, 
ifelse(Hb$AnemiaGroups == 5 & Hb$Hb_AltAdj < 13.0 & Hb$Hb_AltAdj >= 11.00, 1, 0))))))
```

```{r}
# Creating Moderate Category
Hb$AnemiaModerate <- 0
Hb$AnemiaModerate <- ifelse(Hb$AnemiaGroups == 0 & Hb$Hb_AltAdj < 10.0 & Hb$Hb_AltAdj >= 7.00, 1, 
ifelse(Hb$AnemiaGroups == 1 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 8.00, 1, 
ifelse(Hb$AnemiaGroups == 2 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 8.00, 1, 
ifelse(Hb$AnemiaGroups == 3 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 8.00, 1, 
ifelse(Hb$AnemiaGroups == 4 & Hb$Hb_AltAdj < 10.0 & Hb$Hb_AltAdj >= 7.00, 1, 
ifelse(Hb$AnemiaGroups == 5 & Hb$Hb_AltAdj < 11.0 & Hb$Hb_AltAdj >= 8.00, 1, 0))))))
```

```{r}
# Creating Severe Category
Hb$AnemiaSevere <- 0
Hb$AnemiaSevere <- ifelse(Hb$AnemiaGroups == 0 & Hb$Hb_AltAdj < 7.0, 1, 
ifelse(Hb$AnemiaGroups == 1 & Hb$Hb_AltAdj < 8.0, 1, 
ifelse(Hb$AnemiaGroups == 2 & Hb$Hb_AltAdj < 8.0, 1, 
ifelse(Hb$AnemiaGroups == 3 & Hb$Hb_AltAdj < 8.0, 1, 
ifelse(Hb$AnemiaGroups == 4 & Hb$Hb_AltAdj < 7.0, 1, 
ifelse(Hb$AnemiaGroups == 5 & Hb$Hb_AltAdj < 8.0, 1, 0))))))

```

# Distributions
```{r}
table(Hb$Anemia)
skewness(Hb$Anemia, na.rm = T)
kurtosis(Hb$Anemia, na.rm = T)
```

```{r}
table(Hb$Sex)
skewness(Hb$Sex, na.rm = T)
kurtosis(Hb$Sex, na.rm = T)
```

```{r}
describe(Hb$Age)
skewness(Hb$Age, na.rm = T)
kurtosis(Hb$Age, na.rm = T)
hist(Hb$Age)
```

```{r}
plot(Hb$Age, Hb$Anemia)
```

# Participants Section
```{r}
#table(Hb$Region, Hb$Community)
min(Hb$Date)
max(Hb$Date)
describe(Hb$Age)
table(Hb$Sex)
table(Hb$Shuar, exclude = NULL)
table(Hb$Region, exclude = NULL)
table(Hb$Sex, Hb$Region)
table(Hb$Pregnant)
```

# Descriptives Table 1 
```{r}
# Overall Anemia rates
mytable <- table(Hb$Anemia, Hb$AnemiaGroups)
mytable
margin.table(mytable, 1)
prop.table(mytable, 2)

# Mild Anemia rates
table1 <- table(Hb$AnemiaMild, Hb$AnemiaGroups)
table1
margin.table(table1, 1)
prop.table(table1, 2)

# Moderate Anemia Rates 
table2 <- table(Hb$AnemiaModerate, Hb$AnemiaGroups)
table2
margin.table(table2, 1)
prop.table(table2, 2)

# Severe Anemis Rates
table3 <- table(Hb$AnemiaSevere, Hb$AnemiaGroups)
table3
margin.table(table3, 1)
prop.table(table3, 2)

# Mean Hb by group
describe.by(Hb$Hb_AltAdj, group = Hb$AnemiaGroups)
```

# Descriptives Table 2 
```{r}
mean(Hb$Hb_AltAdj[which(Hb$AnemiaMild == 1)])
mean(Hb$Hb_AltAdj[which(Hb$AnemiaModerate == 1)])
mean(Hb$Hb_AltAdj[which(Hb$AnemiaSevere == 1)])

Hb$Children <- ifelse(Hb$AnemiaGroups == 0 | Hb$AnemiaGroups == 1 | Hb$AnemiaGroups == 2, 1, 0)
Hb$Women <- ifelse(Hb$AnemiaGroups == 3 | Hb$AnemiaGroups == 4, 1, 0)
Hb$Men <- ifelse(Hb$AnemiaGroups == 5, 1, 0)

mean(Hb$Hb_AltAdj[which(Hb$Children == 1 & Hb$Region == 1)], na.rm=TRUE)
mean(Hb$Hb_AltAdj[which(Hb$Children == 1 & Hb$Region == 0)], na.rm=TRUE)
mean(Hb$Hb_AltAdj[which(Hb$Women == 1 & Hb$Region == 1)], na.rm=TRUE)
mean(Hb$Hb_AltAdj[which(Hb$Women == 1 & Hb$Region == 0)], na.rm=TRUE)
mean(Hb$Hb_AltAdj[which(Hb$AnemiaGroups == 5 & Hb$Region == 1)], na.rm=TRUE)
mean(Hb$Hb_AltAdj[which(Hb$AnemiaGroups == 5 & Hb$Region == 0)], na.rm=TRUE)

table(Hb$Anemia, Hb$Children, Hb$Region )
table(Hb$Anemia, Hb$Women, Hb$Region)
table(Hb$Anemia, Hb$Men, Hb$Region)
table(Hb$Anemia, Hb$AnemiaGroups, Hb$Region)

table(Hb$Anemia, Hb$Children)
table(Hb$Anemia, Hb$Children, Hb$Sex)
table(Hb$Anemia, Hb$Women)
table(Hb$Anemia, Hb$Men)
table(Hb$Anemia, Hb$AnemiaGroups)
```



# Inferentials: Linear

```{r}
# Preparing data 
Hb$Anemia <- factor(Hb$Anemia)
Hb_glmer  <- filter(Hb, !is.na(Age), !is.na(Sex), !is.na(Community), !is.na(Anemia), !is.na(Year)) 
Hb_glmer$Community <- factor(Hb_glmer$Community)

# Controlling for community and year
## Check to make sure there are enough groups
unique(Hb$Year)

## Anemia and sex

## Adults 15+ 
logit_sex_1  <- glmer(formula = Anemia ~ Sex + Pregnant + (1|Community) + (1|Year), data = Hb_glmer, family = "binomial", subset = AgeGroups == 3)
summary(logit_sex_1)
icc(logit_sex_1)
plogis(0.8059)

## Children Combined
logit_sex_2  <- glmer(formula = Anemia ~ Sex + (1|Community) + (1|Year), data = Hb_glmer, family = "binomial", subset = AgeGroups != 3)
summary(logit_sex_2)
icc(logit_sex_2)
plogis(-0.3458)

```

# Inferentials: GAM
```{r}
# Data prep for GAM
Hb_gam <- as.data.frame(Hb_glmer)
Hb_gam$Community <- as.numeric(factor(Hb_gam$Community))
Hb_gam <- filter(Hb_gam, !is.na(Age), !is.na(Sex), !is.na(Community), !is.na(Anemia), !is.na(Year))

# GAM Model
gam_age <- gamm(Anemia ~ s(Age) + Sex + Pregnant, data = Hb_gam, family = binomial, method = "REML", random=list(Community=~1, Year=~1), na.action = na.omit)

## Model Results
summary(gam_age)
summary(gam_age$gam)
summary(gam_age$lme)

## Assumptions checks
gam.check(gam_age$gam)
concurvity(gam_age$gam)

gam <- gam_age$gam
lme <- gam_age$lme
# Plot log odds

plot(gam, all.terms=F)
vis.gam(gam, view =c("Age", "Sex"))
mean(gam$model$Year)

# PLOT
  # Create new datatable from model
testdata = data.frame(Age=Hb_gam$Age, Year=mean(gam$model$Year), Community=mean(gam$model$Community), Sex=mean(gam$model$Sex), Pregnant=mean(gam$model$Pregnant)) 

# Predict to get fit values and se
p<-predict(gam_age$gam,se.fit=TRUE,newdata=testdata)

# Basic plot of fit
plot(Hb_gam$Age, p$fit)

# Add original data to fit values
p2 <- do.call("cbind", c(testdata, p))
p2 <- data.frame(p2)

# Transform from log using exp
p3 <- mutate(p2, exp_fit = exp(fit), exp_se = exp(se.fit), lower = exp_fit-exp_se, upper = exp_fit+exp_se)

# Plot
ggplot(p3, aes(Age, exp_fit)) +
  geom_line() +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
  labs(y = 'Anemia Odds Ratio', x = "Age") +
  theme_apa()

# Finding Inflection points
p4 <- p3[order(p3$Age), ]
infl <- c(FALSE, diff(diff(p4$fit)>0)!=0)
infl2 <- c(FALSE, infl)
infl3 <- cbind(p4, infl2)
infl4 <- filter(infl3, infl2 == TRUE)

infl2 <- c(FALSE, diff(diff(p4$fit)<0)!=0)
infl2_2 <- c(FALSE, infl2)
infl2_3 <- cbind(p4, infl2_2)
infl2_4 <- filter(infl2_3, infl2_2 == TRUE)
```
# Data Manipulation - Create BMI (RandR2)
```{r}
#Convert height to m from cm
Hb$Height <- Hb$Height/100

# Calculate BMI 
Hb$BMI <- Hb$Weight/(Hb$Height^2)
Hb <- filter(Hb, BMI < 185 | is.na(BMI))
```

```{r}
# (RandR1)
describe(Hb$BMI)
skewness(Hb$BMI, na.rm = T)
kurtosis(Hb$BMI, na.rm = T)
hist(Hb$BMI)
```


# Inferentials regional comparison for revision I (RandR1)
```{r}
Hb_comp <- Hb %>%
  filter(!is.na(Hb_AltAdj), !is.na(Region)) %>%
  group_by(Region) %>%
  summarize(AvgHb = mean(Hb_AltAdj), N = n()) %>%
  print
print(Hb_comp)

glmerRegion <- glmer(Anemia ~ Region + (1|AnemiaGroups) + (1|Year), data = Hb, family = "binomial") 
summary(glmerRegion)
plot(glmerRegion)
qqnorm(resid(glmerRegion))
qqline(resid(glmerRegion))
icc(glmerRegion)
plogis(-0.1351)
```

# Inferentials BMI association for revision I (RandR1)
```{r}
glmerBMI <- glmer(Anemia ~ BMI + (1|AnemiaGroups) + (1|Community) + (1|Year), data = Hb, family = "binomial") 
summary(glmerBMI)
plot(glmerBMI)
qqnorm(resid(glmerBMI))
qqline(resid(glmerBMI))
icc(glmerBMI)
plogis(-0.10213)
```

